---
title: "Call Reports: Reproducible Bank-Level Panel Construction"
author: "Jose Camas Garrdiow"
date: "2024-06-14"
output: 
  html_document:
    theme: readable
    number_sections: true
    df_print: paged
    code_folding: hide
    highlight: tango
    toc: true
    toc_depth: 3
    toc_float: true
    toc_collapsed: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(readr)
library(readxl)
```



# Introduction

This research notebook documents a bank-level panel construction workflow based on U.S. regulatory Call Reports. It is included as a portfolio reconstruction of research assistance performed at IESE Business School. The repository does not redistribute WRDS extracts or any licensed input data; users must supply the source files described below.

The data used in Chakraborty et al. (2020) are cleaned and prepared for analysis. The original extract was obtained through the _Wharton Research Data Services (WRDS)_ platform, specifically from the `SAS Studio` portal, and corresponds to the _Call Reports_ of the _Federal Deposit Insurance Corporation (FDIC)_.

The original source and the replication code was taken from [Phillipp Schnabl's NYU personal website](https://pages.stern.nyu.edu/~pschnabl/), as it is explained [here](https://pages.stern.nyu.edu/~pschnabl/data/data_callreport.htm). The code was later modified to retrieve the raw data from 2005Q4 to 2012Q4, as it is shown in the following chunk:


```{sas, eval=FALSE}
/* SAS Code STARTS HERE */


options linesize=250;
libname mylib "~/"; 

/*This process loops through all call report files and selects the variables*/
%macro loop(start_month=, stop_month=);
    %local month;
    %do month=&start_month %to &stop_month %by 3;
        %put Month: &month;

        data call&month (keep= RSSD9001 RSSD9010 RSSD9348 RSSD9999 RSSD9048
        RCFD2170
        RCON2170
        
        RCFD3210	/* equity*/
        RCON3210
        
        RCFD8639 /* available until 200903 */
        RCON8639
       
        RCFDG379 /* available since 200906 - Residential mortgage pass-through securities issued or guaranteed by FNMA, FHLMC,
or GNM */ 
        RCONG379
        
        RCFDG380 /* available since 200906 - Other residential MBS issued or guaranteed by U.S. Government
agencies or sponsored agencies */ 
        RCONG380
        
        RCFDG381 /* available since 200906 -  All other residential M */ 
        RCONG381
        RCFDK197 /*  available since 201103 - Commercial MBS issued or guaranteed by U.S. Government agencies or sponsored
agencies */
        RCONK197
        RCFDK198 /*available since 201103 -  All other commercial MBS*/
        RCONK198
        RCFD8641 /* available until 201012 */
        RCON8641
        
        RCFD0211
        RCON0211
        RCFD1287
        RCON1287
        RCON3531
        RCFD1289
        RCON1289
        RCFD1294 
        RCON1294
        RCFD1293
        RCON1293
        RCFD1298
        RCON1298
        RCON3532
        RCFD1766
        RCON1766
        RCFDF614
        
        RCONF614
        
        RIAD4012
        RIAD4170 /* available until 201012 */
        
        RCFD2200
        RCON2200
        
        RCFD0010
        RCON0010
        
        RCFD2122
        RCON2122
        
        RCFD2210
        RCON2210
        
        RIAD4340 /* netincome */
        
        RIADB493
        RCON2746
        RCFD2746
        RCON1564
        RCON1975
        RCON1480
        RCON3816

        /* added variables 31.08 */
        rcon1763
        rcon1764
        rcon2008
        rconb538
        rconb539
        rcon2011
        rconk137
        rconk207
        rcon2182
        rconf162
        rcon2183
        rconf163
        rcon4301
        rcon4340
        rcon4460
        rcon8274
        rcoa8274
        rcon5479
        rcon3247
        rcon7204
        rcon3000
        rcon3284
        rcon1415
        RCONF158
        RCONF159
        RCON1420
        RCON1460
        RCON1797
        RCON1798
        RCON5367
        RCON5368
        RCON2081
        RCON1545
        rcon0383
        rcon0384
        rcon0387
        rcon3533
        rcon9802
        rcop2200
        rcon4073
        rcon4074
        rcon4079
        rcon4107
        rcon2309
        rssd9005
        rssd9007
        rssd9008
        rssd9016
        rssd9017
        rssd9030
        rssd9031
        rssd9032
        rssd9037
        rssd9038
        rssd9039
        rssd9040
        rssd9042
        rssd9043
        rssd9044
        rssd9045
        rssd9047
        rssd9046
        rssd9049
        rssd9051
        rssd9052
        rssd9053
        rssd9054
        rssd9055
        rssd9056
        rssd9057
        rssd9059
        rssd9061
        rssd9074
        rssd9101
        rssd9130
        rssd9950
        rssd9150
        rssd9161
        rssd9200
        rssd9210
        rssd9213
        rssd9217
        rssd9241
        rssd9331
        rssd9349
        rssd9375
        rssd9425
        rssd9397
        );
        set bank.call&month;
    run;

    %if %substr(&month,5,2)=12 %then %let month=%eval(&month+88);
    %end;
%mend loop;
%loop(start_month=199012,stop_month=202012)

data db;
     set call:; /*Merging all datasets starting by bhcf in one called temp */
     /*rename RSSD9999=date RSSD9001 = rssdid RSSD9010 = name RSSD9348 = bhcid RSSD9048 = chartype;*/
run;

data db(compress=yes);
      set db; 
run;

proc export data=db 
    outfile="~/chakraborty2022_rawdb_2.csv" 
    dbms=csv 
    replace;
run;

```



After that, the file `chakraborty2022_rawdb_2.csv` was downloaded and saved in the `DataWRDS/ReplicationChakraborty2020` folder. 

The following string reads the data and shows the first rows of the dataset:

```{r paths}
project_dir <- normalizePath(
  Sys.getenv("CALL_REPORTS_PROJECT_DIR", unset = "."),
  mustWork = FALSE
)
data_dir <- file.path(project_dir, "data", "raw")
output_dir <- file.path(project_dir, "outputs")
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
```

```{r}
calls_original <- read.csv(
  file.path(data_dir, "chakraborty2022_rawdb_2.csv")
)
head(calls_original)
```

Note that the information is expressed in thousands of dollars.

# Data cleaning

The following chunk of code cleans the data and prepares it for the analysis. The main steps are:

1. Convert the RSSD9999 variable to a date format.
2. Arrange the data by RSSD9001 and date.
3. Create a variable `quarter` that indicates the quarter of the year.
4. Create a variable `delta` that indicates the difference between the quarter and the first quarter of the year.
5. Create a variable `diff_delta` that indicates the difference between the current `delta` and the previous `delta`.
6. Replace missing values in the `RCFD` series with the `RCON` series. (This is because starting from 20110331, banks without foreign offices do not report RCFD series)
7. Create quarterly variables for the income statement due to those variables (`RIAD`) are reported from year to date (on a cumulative basis, that is earnings in Q2 are the sum of Q1 and Q2 earnings, etc.)
8. Finally, according to the authors, there are some cases were the same data was reported for the same bank in the same quarter. In those cases, the observation with the less amount of assets is dropped.

All those steps are performed in the following chunk of code:

```{r}
calls <- calls_original |> 
  mutate(date = lubridate::ymd(RSSD9999)) |> #create date
  arrange(RSSD9001,date) |>
  mutate(
    quarter = lubridate::quarter(date),
    delta = quarter*25,
    diff_delta = delta - dplyr::lag(delta),
    # Starting from 20110331, banks without foreign offices do not report RCFD series. 
    # This process replaces the missing RCFD values with RCON values
    (across(starts_with("RCFD"), 
            ~case_when(
              is.na(.) ~  get(paste0("RCON", sub("RCFD", "", cur_column()))),
              TRUE ~ .),
            .names = "{.col}")),
    (across(starts_with('RIAD'),
            ~ case_when(
              quarter == 1 ~ .,
              quarter != 1 & diff_delta == 25 ~ . - dplyr::lag(.),
              TRUE ~ NA_real_),
            .names = "{.col}"))
  ) |> 
  # There are some banks that were repeated, the following code removes the duplicates
  arrange(RSSD9001, date, desc(RCFD2170)) |>
  distinct(RSSD9001,date, .keep_all = TRUE)
```


## 1. VISUALIZATION. COMPARISON DATASET-OTHER SOURCES


### Assets

First, let's plot the assets over time:

```{r}
calls |>
  group_by(date) |>
  summarise(RCFD2170 = sum(RCFD2170/1000000000, na.rm = T)) |>
  ggplot(aes(x = date, y = RCFD2170)) +
  geom_line(linewidth= 1) +
  labs(title = "Assets over time",
       x = "Date",
       y = "Assets (in trillions of dollars)")

```
Even though the plot is very clear, there is a huge drop in the assets in 2013. In the following plot the information is broken down by type of entity:


```{r 'assetsbytype', fig.width=14, fig.height=9, message=FALSE}
calls |> 
  mutate(entity_type = case_when(
    RSSD9048 == 0 ~ 'NA or not applicable (for branches)',
    RSSD9048 == 100 ~ 'Government',
    RSSD9048 == 110 ~ 'Government Agency ',
    RSSD9048 == 120 ~ 'Overseeing Regulatory Body',
    RSSD9048 == 130 ~ 'U.S. Central Bank or Liquidity Agency',
    RSSD9048 == 200 ~ 'Commercial Bank ',
    RSSD9048 == 250 ~ 'Nondeposit Trust Company',
    RSSD9048 == 300 ~ 'Savings Bank',
    RSSD9048 == 310 ~ 'Savings & Loan Association',
    RSSD9048 == 320 ~ 'Cooperative Bank',
    RSSD9048 == 330 ~ 'Credit Union (excluding Corporates)',
    RSSD9048 == 340 ~ 'Industrial Bank ',
    RSSD9048 == 350 ~ 'Article XII New York Investment Company',
    RSSD9048 == 370 ~ 'Corporate Credit Union',
    RSSD9048 == 400 ~ 'Edge or Agreement Corporation',
    RSSD9048 == 500 ~ 'Holding Company only, not itself any other charter type (not necessarily a BHC)',
    RSSD9048 == 550 ~ 'Insurance Broker or Agent and/or Insurance Company (includes insurance underwriting)',
    RSSD9048 == 610 ~ 'Employee Stock Ownership Plan/Trust (ESOP or ESOT)',
    RSSD9048 == 700 ~ 'Securities Broker and/or Dealer (includes securities underwriting)',
    RSSD9048 == 710 ~ 'Utility Company or Electric Power Co-generator',
    RSSD9048 == 720 ~ 'Other Non Depository Institution',
    RSSD9048 == 800 ~ 'Pseudo Entity / Accounting',
    RSSD9048 == 810 ~ 'Currency, Verification, Counting and Sorting Machine',
    RSSD9048 == 820 ~ 'Direct Mint Shipment Site, not itself any other entity type',
    RSSD9048 == 900 ~ 'Individual or Group of Individuals',
    TRUE ~ 'Other'
  )) |> 
  select(date,entity_type,RSSD9001,RCFD2170 ) |> 
#  filter(RSSD9001 == 1511) |> 
  group_by(date,entity_type) |>
  summarise(across(starts_with("RCFD"), list('sum'= ~sum(., na.rm=T)))) |>
  ggplot(aes(x=date, y=RCFD2170_sum/1000000000)) +
  #geom_rect(aes(xmin = ymd('2007-11-27'), xmax = ymd('2009-05-27'), ymin = -Inf, ymax = Inf), fill = 'gray90', alpha = 0.2, inherit.aes = FALSE) +
  geom_line(linewidth = 1) + 
  #theme_minimal()+
  facet_wrap(~entity_type, scales = 'free_y')+
  #scale_color_brewer(palette = 'Set2') +
  labs(title='Time series of total assets by entity type',
       x='Date',
       y='Value (trillions)') +
  scale_x_date(date_labels = '%Y-%m', date_breaks = '1 year')+
  theme(legend.position = 'bottom', 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
```

```{r}
library(dplyr)
library(ggplot2)
library(data.table)
library(lubridate)

commercial_banks_plot <- calls %>%
  filter(RSSD9048 == 200) %>%
  group_by(date) %>%
  summarise(assets = sum(RCFD2170 / 1000000000, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = assets)) +
  geom_line(linewidth = 1, color = "#2c7bb6") +  # Adjusted color to match FRED style
  labs(title = 'Total Assets, All Commercial Banks',
       x = NULL,
       y = 'Billions of U.S. Dollars') +
  scale_x_date(date_labels = '%Y-%m', date_breaks = '1 year', expand = c(0, 0)) +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "grey80"),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.title = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )

print(commercial_banks_plot)
```

It seems that the drop in assets is mainly due to the lack of reporting of _Edge or Agreement Corporation_ and _NA or not applicable (for branches)_ entities, even tough there is a sharp increase in the _Saving Bank_ and _Savings & Loan Association_ entities.

Also, as it can be seen from the image below, the assets of the _Commercial Bank_ entities match the data published by the St. Louis Federal Reserve (FRED) [here](https://fred.stlouisfed.org/series/USTAST).


### Equity

The following plot shows the equity over time:

```{r}
calls |>
  group_by(date) |>
  summarise(RCFD3210 = sum(RCFD3210/1000000000, na.rm = T)) |>
  ggplot(aes(x = date, y = RCFD3210)) +
  geom_line(linewidth= 1) +
  labs(title = "Equity over time",
       x = "Date",
       y = "Equity (in trillions of dollars)")

```

Now, let's break down the equity by entity type:

```{r 'equitybytype', fig.width=14, fig.height=9, message=FALSE}
calls |> 
  mutate(entity_type = case_when(
    RSSD9048 == 0 ~ 'NA or not applicable (for branches)',
    RSSD9048 == 100 ~ 'Government',
    RSSD9048 == 110 ~ 'Government Agency ',
    RSSD9048 == 120 ~ 'Overseeing Regulatory Body',
    RSSD9048 == 130 ~ 'U.S. Central Bank or Liquidity Agency',
    RSSD9048 == 200 ~ 'Commercial Bank ',
    RSSD9048 == 250 ~ 'Nondeposit Trust Company',
    RSSD9048 == 300 ~ 'Savings Bank',
    RSSD9048 == 310 ~ 'Savings & Loan Association',
    RSSD9048 == 320 ~ 'Cooperative Bank',
    RSSD9048 == 330 ~ 'Credit Union (excluding Corporates)',
    RSSD9048 == 340 ~ 'Industrial Bank ',
    RSSD9048 == 350 ~ 'Article XII New York Investment Company',
    RSSD9048 == 370 ~ 'Corporate Credit Union',
    RSSD9048 == 400 ~ 'Edge or Agreement Corporation',
    RSSD9048 == 500 ~ 'Holding Company only, not itself any other charter type (not necessarily a BHC)',
    RSSD9048 == 550 ~ 'Insurance Broker or Agent and/or Insurance Company (includes insurance underwriting)',
    RSSD9048 == 610 ~ 'Employee Stock Ownership Plan/Trust (ESOP or ESOT)',
    RSSD9048 == 700 ~ 'Securities Broker and/or Dealer (includes securities underwriting)',
    RSSD9048 == 710 ~ 'Utility Company or Electric Power Co-generator',
    RSSD9048 == 720 ~ 'Other Non Depository Institution',
    RSSD9048 == 800 ~ 'Pseudo Entity / Accounting',
    RSSD9048 == 810 ~ 'Currency, Verification, Counting and Sorting Machine',
    RSSD9048 == 820 ~ 'Direct Mint Shipment Site, not itself any other entity type',
    RSSD9048 == 900 ~ 'Individual or Group of Individuals',
    TRUE ~ 'Other'
  )) |> 
  select(date,entity_type,RSSD9001,RCFD3210 ) |>
  group_by(date,entity_type) |>
  summarise(across(starts_with("RCFD"), list('sum'= ~sum(., na.rm=T)))) |>
  ggplot(aes(x=date, y=RCFD3210_sum/1000000000)) +
  #geom_rect(aes(xmin = ymd('2007-11-27'), xmax = ymd('2009-05-27'), ymin = -Inf, ymax = Inf), fill = 'gray90', alpha = 0.2, inherit.aes = FALSE) +
  geom_line(linewidth = 1) +
  facet_wrap(~entity_type, scales = 'free_y')+
  labs(title='Time series of equity by entity type',
       x='Date',
       y='Value (trillions of USD)') +
  scale_x_date(date_labels = '%Y-%m', date_breaks = '1 year')+
  theme(legend.position = 'bottom', 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

The same pattern is observed in the equity time series, with the _Edge or Agreement Corporation_ entities not reporting their equity.


### Net income

The following plot shows the net income over time for the raw data:

```{r}
calls_original |> 
  mutate(date = lubridate::ymd(RSSD9999)) |> 
  group_by(date) |>
  summarise(RIAD4340 = sum(RIAD4340/1000000000, na.rm = T)) |>
  ggplot(aes(x = date, y = RIAD4340)) +
  geom_line(linewidth= 1) +
  labs(title = "Net income over time",
       x = "Date",
       y = "Net income (in billions of dollars)")

```

The previous plot shows the year to date net income for the raw data. As the income is cumulative, the plot shows a clear increasing trend over the year. However, the following plot shows the net income over time for the cleaned data:

```{r}
calls |>
  filter(date >= dmy(01012006)) |> 
  group_by(date) |>
  summarise(RIAD4340 = sum(RIAD4340/1000000000, na.rm = T)) |>
  ggplot(aes(x = date, y = RIAD4340)) +
  geom_line(linewidth= 1) +
  labs(title = "Net income over time",
       subtitle = "Net income for 2005q4 was removed for plotting because it showed a cumulative value.",
       x = "Date",
       y = "Net income (in billions of dollars)")

```

Now the series is plotted by entity type:

```{r 'netincomebytype', fig.width=14, fig.height=9, message=FALSE}

calls |> 
  filter(date >= dmy(01012006)) |>
  mutate(entity_type = case_when(
    RSSD9048 == 0 ~ 'NA or not applicable (for branches)',
    RSSD9048 == 100 ~ 'Government',
    RSSD9048 == 110 ~ 'Government Agency ',
    RSSD9048 == 120 ~ 'Overseeing Regulatory Body',
    RSSD9048 == 130 ~ 'U.S. Central Bank or Liquidity Agency',
    RSSD9048 == 200 ~ 'Commercial Bank ',
    RSSD9048 == 250 ~ 'Nondeposit Trust Company',
    RSSD9048 == 300 ~ 'Savings Bank',
    RSSD9048 == 310 ~ 'Savings & Loan Association',
    RSSD9048 == 320 ~ 'Cooperative Bank',
    RSSD9048 == 330 ~ 'Credit Union (excluding Corporates)',
    RSSD9048 == 340 ~ 'Industrial Bank ',
    RSSD9048 == 350 ~ 'Article XII New York Investment Company',
    RSSD9048 == 370 ~ 'Corporate Credit Union',
    RSSD9048 == 400 ~ 'Edge or Agreement Corporation',
    RSSD9048 == 500 ~ 'Holding Company only, not itself any other charter type (not necessarily a BHC)',
    RSSD9048 == 550 ~ 'Insurance Broker or Agent and/or Insurance Company (includes insurance underwriting)',
    RSSD9048 == 610 ~ 'Employee Stock Ownership Plan/Trust (ESOP or ESOT)',
    RSSD9048 == 700 ~ 'Securities Broker and/or Dealer (includes securities underwriting)',
    RSSD9048 == 710 ~ 'Utility Company or Electric Power Co-generator',
    RSSD9048 == 720 ~ 'Other Non Depository Institution',
    RSSD9048 == 800 ~ 'Pseudo Entity / Accounting',
    RSSD9048 == 810 ~ 'Currency, Verification, Counting and Sorting Machine',
    RSSD9048 == 820 ~ 'Direct Mint Shipment Site, not itself any other entity type',
    RSSD9048 == 900 ~ 'Individual or Group of Individuals',
    TRUE ~ 'Other'
  )) |> 
  select(date,entity_type,RIAD4340 ) |>
  group_by(date,entity_type) |>
  summarise(across(starts_with("RIAD"), list('sum'= ~sum(., na.rm=T)))) |>
  ggplot(aes(x=date, y=RIAD4340_sum/1000000)) +
  geom_line(linewidth = 1) +
  facet_wrap(~entity_type, scales = 'free_y')+
  labs(title='Time series of net income by entity type',
       x='Date',
       y='Value (billions of USD)') +
  scale_x_date(date_labels = '%Y-%m', date_breaks = '1 year')+
  theme(legend.position = 'bottom', 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

___________________________________________________________________________________________________________________________________________________________________


## F2. FINAL DATASET

We took into account the dataset from Chakraborty but in that dataset there are not granular variables for deposit. In order to have a deeper understanding of sourcing, we merge these data with a dataset used in Schnabl's papers "The deposits channel of monetary policy" and "Banking on Deposits: Maturity Transformation without Interest Rate Risk", following the instructions from the website https://pages.stern.nyu.edu/~pschnabl/data/data_callreport.htm. Both dataset are retrieved from the same WRDS source, but Schnabl's data are wider in terms of interval and include more variables. The following is the code that was used to create the final dataset, which can be eventually replicated, whenever necessary with other variables. 


# Step 1: Convert the date column in callreports_1976_2020_WRDS to date format yyyy-mm-dd
``{r}
callreports <- read.csv(
  file.path(data_dir, "callreports_1976_2020_WRDS.csv")
)
library(data.table)

setDT(calls)
setDT(callreports)

callreports[, date := as.Date(as.character(date), format = "%Y%m%d")]

# Step 2: Select the relevant variables from callreports_1976_2020_WRDS
selected_columns <- c("date", "rssdid", "brokereddeple100k", "brokereddepeq100k", "qavgsavdep", 
                      "qavgtimedepge100k", "qavgtimedepge250k", "qavgcdge100k", 
                      "qavgtimedeple100k", "qavgtimedeple250k", "qavgfordep", 
                      "timedeple100k_less_3m", "timedeple250k_less_3m", "timedeple100k_3m_1y", 
                      "timedeple250k_3m_1y", "timedeple100k_1y_3y", "timedeple250k_1y_3y", 
                      "timedeple100k_over_3y", "timedeple250k_over_3y", "timedeple100k_less_1y", 
                      "timedeple250k_less_1y", "timedepge100k_less_3m", "timedepge250k_less_3m", 
                      "timedepge100k_3m_1y", "timedepge250k_3m_1y", "timedepge100k_1y_3y", 
                      "timedepge250k_1y_3y", "timedepge100k_over_3y", "timedepge250k_over_3y")

# Additional columns
additional_columns <- c(
  "reloans", "persloans", "agloans",
  "loansnet", "tradingassets",
  "securitiesheldtomaturity", "securitiesavailableforsale", "deposits", "foreigndep", "nonintbeardep", 
  "intbeardep", "intbearfordep",
  "timedepuninsured",
  "intexp", "intincnet", "nonintinc", "domdepservicecharges", "nonintexp",
  "numemployees", "intexpalldep",
  "intexpdomdep", "intexpsubordinated", "intanddivincsecurities", "loanleaselossprovision", 
  "intexptimedepge100k", "intexptimedeple100k", "intexptimedeple250k", "intexptimedepge250k", 
  "intexpcdge100k", "intexptimedep",
  "intincloans", "intincreloans",
  "resloans_less_3m", "resloans_3m_1y", "resloans_1y_3y", "resloans_3y_5y", "resloans_5y_15y", 
  "resloans_over_15y"
)

# Combine and remove duplicates
selected_columns <- unique(c(selected_columns, additional_columns))
```


``{r}
callreports_selected <- callreports[, ..selected_columns]
```

``{r}
# Convert the date columns to Date format
calls$date <- as.Date(calls$date)
callreports_selected$date <- as.Date(callreports_selected$date)

merged_data <- calls %>%
  inner_join(callreports_selected, by = c("date" = "date", "RSSD9001" = "rssdid"))

```

``{r}
write.csv(
  merged_data,
  file.path(output_dir, "final_dataset.csv"),
  row.names = FALSE
)
```

___________________________________________________________________________________________________________________________________________________________________

## 3. CHACKRABORTY CONSISTENCY 

### Replicating summary statistics

The table it is aimed to replicate is Table 1 from Chackraborty (2020), whose variables construction method is defined in Table A.1 from the appendix, which lists also the data sources.

To create the necessary variables, we started by loading the required dataset and libraries. We then used the mutate function from the dplyr package to generate each variable according to the specified formulas. This involved calculating values like MBS holdings, securities holdings, U.S. government securities holdings, C&I loan growth, change in C&I loan profitability, bank size, bank equity ratio, bank net income, bank cost of deposits, and bank demand deposits. Each new variable was scaled appropriately and based on the provided dataset columns.


```{r}
# Ensure the date column is in date format and filter the data
calls <- calls %>%
    filter(date >= as.Date("1990-12-31") & date <= as.Date("2020-12-31"))

calls <- calls %>%
  mutate(across(everything(), ~replace_na(., 0)))

calls <- calls %>%
  mutate(
    MBS_Holdings = ((RCFD8639 + RCFDG379 + RCFDG380 + RCFDG381 + RCFDK197 + RCFDK198) / RCFD2170) * 100,
    Securities_Holdings = ((RCFD8641 - RCFD8639) / RCFD2170) * 100,
    US_Gov_Securities_Holdings = ((RCFD0211 + RCFD1287 + RCON3531 + RCFD1289 + RCFD1294 + RCFD1293 + RCFD1298 + RCON3532) / RCFD2170) * 100,
    C_I_Loan_Growth = (((RCFD1766 + RCFDF614) / lag(RCFD1766 + RCFDF614)) - 1) * 100,
    Change_in_C_I_Loan_Profitability = (((RIAD4012 / RCFD1766) - lag(RIAD4012 / RCFD1766)) * 100),
    Bank_Size = log(RCFD2170),
    Bank_Equity_Ratio = ((RCFD3210 / RCFD2170) * 100),
    Bank_Net_Income = ((RIAD4340 / RCFD2170) * 100),
    Bank_Cost_of_Deposits = ((RIAD4170 / RCFD2200) * 100),
    Bank_Demand_Deposits = ((RCFD2210 / RCFD2170) * 100)
  )

# sum(!is.na(calls$MBS_Holdings))

# sum(!is.na(calls$RCFD2170) & calls$RCFD2170 != 0)

```

For calculating summary statistics, we utilized the summarise function from the dplyr package. This function helped us compute the mean, standard deviation, 25th percentile, median, and 75th percentile for each of the newly created variables. We ensured that all calculations excluded any missing values by using the na.rm = TRUE parameter. These summary statistics provide a comprehensive overview of the distribution and central tendencies of the bank variables, which are essential for further analysis and reporting.

```{r}

# Load necessary libraries
library(dplyr)
library(purrr)
library(knitr)
library(kableExtra)

# Define a Winsorizing function
winsorize <- function(x, probs = c(0.02, 0.963)) {
  quantiles <- quantile(x, probs, na.rm = TRUE)
  x[x < quantiles[1]] <- quantiles[1]
  x[x > quantiles[2]] <- quantiles[2]
  return(x)
}

# Ensure the date column is in date format and filter the data
calls <- calls %>%
    filter(date >= as.Date("1990-12-31") & date <= as.Date("2020-12-31"))

calls <- calls %>%
  mutate(across(everything(), ~replace_na(., 0)))

# Create new variables
calls <- calls %>%
  mutate(
    MBS_Holdings = ((RCFD8639 + RCFDG379 + RCFDG380 + RCFDG381 + RCFDK197 + RCFDK198) / RCFD2170) * 100,
    Securities_Holdings = ((RCFD8641 - RCFD8639) / RCFD2170) * 100,
    US_Gov_Securities_Holdings = ((RCFD0211 + RCFD1287 + RCON3531 + RCFD1289 + RCFD1294 + RCFD1293 + RCFD1298 + RCON3532) / RCFD2170) * 100,
    C_I_Loan_Growth = (((RCFD1766 + RCFDF614) / lag(RCFD1766 + RCFDF614)) - 1) * 100,
    Change_in_C_I_Loan_Profitability = (((RIAD4012 / RCFD1766) - lag(RIAD4012 / RCFD1766)) * 100),
    Bank_Size = log(RCFD2170),
    Bank_Equity_Ratio = ((RCFD3210 / RCFD2170) * 100),
    Bank_Net_Income = ((RIAD4340 / RCFD2170) * 100),
    Bank_Cost_of_Deposits = ((RIAD4170 / RCFD2200) * 100),
    Bank_Demand_Deposits = ((RCFD2210 / RCFD2170) * 100)
  )

# Windsorize the variables except Bank_Size
calls <- calls %>%
  mutate(
    MBS_Holdings = winsorize(MBS_Holdings),
    Securities_Holdings = winsorize(Securities_Holdings),
    US_Gov_Securities_Holdings = winsorize(US_Gov_Securities_Holdings),
    C_I_Loan_Growth = winsorize(C_I_Loan_Growth),
    Change_in_C_I_Loan_Profitability = winsorize(Change_in_C_I_Loan_Profitability),
    Bank_Equity_Ratio = winsorize(Bank_Equity_Ratio),
    Bank_Net_Income = winsorize(Bank_Net_Income),
    Bank_Cost_of_Deposits = winsorize(Bank_Cost_of_Deposits),
    Bank_Demand_Deposits = winsorize(Bank_Demand_Deposits)
  )

# Define a function to compute summary statistics excluding zeros and infinite values
summary_stats_function <- function(x) {
  valid_x <- x[x != 0 & is.finite(x)]
  if (length(valid_x) > 0) {
    data.frame(
      mean = mean(valid_x, na.rm = TRUE),
      sd = sd(valid_x, na.rm = TRUE),
      p25 = quantile(valid_x, 0.25, na.rm = TRUE),
      median = median(valid_x, na.rm = TRUE),
      p75 = quantile(valid_x, 0.75, na.rm = TRUE),
      n = length(valid_x)
    )
  } else {
    data.frame(
      mean = NA,
      sd = NA,
      p25 = NA,
      median = NA,
      p75 = NA,
      n = 0
    )
  }
}

# Compute the summary statistics for the specified variables
summary_stats <- bind_rows(
  MBS_Holdings = summary_stats_function(calls$MBS_Holdings),
  Securities_Holdings = summary_stats_function(calls$Securities_Holdings),
  US_Gov_Securities_Holdings = summary_stats_function(calls$US_Gov_Securities_Holdings),
  C_I_Loan_Growth = summary_stats_function(calls$C_I_Loan_Growth),
  Change_in_C_I_Loan_Profitability = summary_stats_function(calls$Change_in_C_I_Loan_Profitability),
  Bank_Size = summary_stats_function(calls$Bank_Size),
  Bank_Equity_Ratio = summary_stats_function(calls$Bank_Equity_Ratio),
  Bank_Net_Income = summary_stats_function(calls$Bank_Net_Income),
  Bank_Cost_of_Deposits = summary_stats_function(calls$Bank_Cost_of_Deposits),
  Bank_Demand_Deposits = summary_stats_function(calls$Bank_Demand_Deposits),
  .id = "Variable"
)

# Rename columns to match desired format
summary_stats <- summary_stats %>%
  rename(
    "Mean" = mean,
    "Standard deviation" = sd,
    "25th percentile" = p25,
    "Median" = median,
    "75th percentile" = p75,
    "Number of observations" = n
  )

# Create the table
kable(summary_stats, "html", escape = FALSE, align = 'c') %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  add_header_above(c(" " = 1, "Summary Statistics" = 7))

print(summary_stats)
```

The only value that is not giving results aligned with those from Chakraborty is related to Change in C&I Loan Profitability. All the others are pretty similar. I winsorized all the variables but Bank Size trying to minimize the problems connected with extreme values that, without winsorization were affecting especially Lean Growth.

____________________________________________________________________________________________________________________________________________

## 4. COMPARING BETWEEN AGGREGATED DATA AND HOLDINGS

In order to retrieve the data for holding banks, I considered the dataset WRDS_HOLDING_BHCK from BANK library on SAS Studio WRDS Cloud. I filtered by the variables that we had already in the chackraborty dataset. 

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(lubridate)
```

### LOADING AND PREPARATION

```{r}
# Read the original calls data from chakraborty raw file
calls_original <- read.csv(
  file.path(data_dir, "chakraborty2022_rawdb_2.csv")
)

# Load the Variables_Holders dataset
variables_holders <- read.csv(
  file.path(data_dir, "Variables_Holdings.csv")
)

# Convert RSSD9999 to a date format and calculate the corresponding fiscal quarter
variables_holders <- variables_holders %>%
  mutate(date = ymd(RSSD9999),                           # Convert RSSD9999 to proper date format
         quarter = floor_date(date, "quarter"))          # Calculate the fiscal quarter

# Load and prepare the calls data by converting date and calculating fiscal quarters
calls <- read.csv(file.path(data_dir, "calls.csv"))
calls <- calls %>%
  mutate(date = ymd(date),
         quarter = floor_date(date, "quarter"))
```

### FILTERING 
```{r, echo=FALSE}
# Filter both datasets for the interval 2005Q4 to 2014Q4
start_date <- ymd("1990-12-31")
end_date <- ymd("2020-12-31")

variables_holders_filtered <- variables_holders %>%
  filter(date >= start_date & date <= end_date)

calls_filtered <- calls %>%
  filter(date >= start_date & date <= end_date)
```


### AGGREGATION 
```{r, echo=FALSE}
# Aggregate data per quarter
variables_holders_quarterly <- variables_holders_filtered %>%
  group_by(quarter) %>%
  summarise(across(starts_with("BHCK"), sum, na.rm = TRUE))

calls_quarterly <- calls_filtered %>%
  group_by(quarter) %>%
  summarise(across(starts_with("RCFD"), sum, na.rm = TRUE),
            across(starts_with("RIAD"), sum, na.rm = TRUE))

# Merge the two aggregated datasets by quarter
merged_data <- merge(calls_quarterly, variables_holders_quarterly, by = "quarter", all = TRUE)
```

### PLOTTING PREPARATION
```{r, echo=FALSE}
# Variable pairs to compare
variable_pairs <- list(
  c("RCFD1287", "BHCK1287", "Fair Value of AFS US Treasury Securities"),
  c("RCFD1289", "BHCK1289", "Amortized Cost of HTM US Govt. Agency & Corp. Obligations (Excl. MBS)"),
  c("RCFD1293", "BHCK1293", "Fair Value of AFS US Govt. Agency & Corp. Obligations (Excl. MBS)"),
  c("RCFD1294", "BHCK1294", "Amortized Cost of HTM US Govt. Agency & Corp. Obligations (Excl. MBS)"),
  c("RCFD1298", "BHCK1298", "Fair Value of AFS US Govt. Agency & Corp. Obligations (Excl. MBS)"),
  c("RCFD1766", "BHCK1766", "Commercial & Industrial Loans")
)

# Function to plot comparison
plot_comparison <- function(call_var, holder_var, title, data) {
  p <- ggplot(data, aes(x = quarter)) +
    geom_line(aes(y = .data[[call_var]], color = "call_report_aggregated"), size = 1.2) +
    geom_line(aes(y = .data[[holder_var]], color = "holding_banks"), size = 1.2) +
    labs(title = title,
         x = "Quarter", y = "Value") +
    scale_color_manual(name = "Source", values = c("call_report_aggregated" = "blue", "holding_banks" = "red")) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
      axis.title = element_text(size = 14),
      axis.text = element_text(size = 12),
      legend.title = element_text(size = 14),
      legend.text = element_text(size = 12)
    )
  
  return(p)
}
```

### PLOTTING EXECUTION

```{r, echo=FALSE}
# Plot all comparisons
for (pair in variable_pairs) {
  call_var <- pair[1]
  holder_var <- pair[2]
  title <- pair[3]
  
  p <- plot_comparison(call_var, holder_var, title, merged_data)
  
  print(p)
}



# Net income Holdings vs aggregated

# Define date range
start_date <- ymd("1990-12-31")
end_date <- ymd("2020-12-31")

# Prepare the data for RIAD4340
riad_data <- calls_original %>%
  mutate(date = lubridate::ymd(RSSD9999)) %>%
  filter(date >= start_date & date <= end_date) %>%
  group_by(date) %>%
  summarise(RIAD4340 = sum(RIAD4340 / 1000000000, na.rm = TRUE))

# Prepare the data for BHCK4340
bhck_data <- variables_holders %>%
  mutate(date = lubridate::ymd(RSSD9999)) %>%
  filter(date >= start_date & date <= end_date) %>%
  group_by(date) %>%
  summarise(BHCK4340 = sum(BHCK4340 / 1000000000, na.rm = TRUE))

# Merge the two datasets by date
combined_data <- merge(riad_data, bhck_data, by = "date", all = TRUE)

# Plot the combined data
ggplot(combined_data, aes(x = date)) +
  geom_line(aes(y = RIAD4340, color = "call_report_aggregated"), size = 1) +
  geom_line(aes(y = BHCK4340, color = "holding_banks"), size = 1) +
  labs(title = "Net Income (loss) over time (in billions)",
       x = "Date",
       y = "Value",
       color = "Variable") +
  scale_color_manual(values = c("call_report_aggregated" = "blue", "holding_banks" = "red")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )

# Net Securitization Holdings vs aggregated

# Prepare the data for RIADB493
riad_data <- calls_original %>%
  mutate(date = lubridate::ymd(RSSD9999)) %>%
  filter(date >= start_date & date <= end_date) %>%
  group_by(date) %>%
  summarise(RIADB493 = sum(RIADB493, na.rm = TRUE))

# Prepare the data for BHCKB493
bhck_data <- variables_holders %>%
  mutate(date = lubridate::ymd(RSSD9999)) %>%
  filter(date >= start_date & date <= end_date) %>%
  group_by(date) %>%
  summarise(BHCKB493 = sum(BHCKB493, na.rm = TRUE))

# Merge the two datasets by date
combined_data <- merge(riad_data, bhck_data, by = "date", all = TRUE)

# Plot the combined data
ggplot(combined_data, aes(x = date)) +
  geom_line(aes(y = RIADB493, color = "call_report_aggregated"), size = 1) +
  geom_line(aes(y = BHCKB493, color = "holding_banks"), size = 1) +
  labs(title = "Net Securitization Income",
       x = "Date",
       y = "Value",
       color = "Variable") +
  scale_color_manual(values = c("call_report_aggregated" = "blue", "holding_banks" = "red")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )
```

In these plots related to variables we could compare with holding banks data we can observe how the trends between holding bank and commercial banks aggergated seem to be consistent. We also observe that plots of data from call report aggregated commercial banks are always a bit higher, which is possible to explain considering that holdings' banks data have a certain threshold under which data are not considered. 

________________________________________________________________________________________________

## 5. AGGREGATING AT TOP HOLDING LEVEL

In order to have aggregated data at top_holding level, we proceeded by creating a dataset that drops rssdid and bhcid, creating a new variable which is called bankid. When bhcid (which in our dataset is identified by RSSD9348) is equal to 0, it means that the bank is already a top holding (i.e. there are no banks that own that bank). When it's different from 0, the value displayed is the rssdid of the top ownwer (top holding). 

```{r}
Final_Dataset <- read.csv(file.path(data_dir, "Final_Dataset.csv"))
```


```{r}
unique_bhcid_count <- Final_Dataset %>%
  summarise(unique_count_bhcid = n_distinct(RSSD9348))

print(unique_bhcid_count)

unique_rssdid_count <- Final_Dataset %>%
  summarise(unique_count_rssdid = n_distinct(RSSD9001))

print(unique_rssdid_count)

```

# Aggregate at top holding level

To create the bankid variable we considered the value in RSSD9348 itself when the columns RSSD9348 is different than 0 (therefore, we directly listed the right top holding); for RSSD9348 = 0 or NA we directly considered the rssdid, because that identifier is already the one of the top holding.

```{r}
# Create the bankid column
data_filtered <- Final_Dataset %>%
  mutate(bankid = if_else(RSSD9348 == 0 | is.na(RSSD9348), RSSD9001, RSSD9348))
```

```{r}
unique_bankid_count <- data_filtered %>%
  summarise(unique_count_bankid = n_distinct(bankid))

print(unique_bankid_count)
```

A bit of cleaning to drop columns rssdid and bhcid, plus the column with the name of the subsidiaries. 

```{r}
data_filtered <- data_filtered %>%
  select(RSSD9001, RSSD9348, RSSD9010, date, bankid, RSSD9048, everything())
```

```{r}
data_filtered <- data_filtered[, -c(1:3,7)]
```

Once all this is done, we can finally aggregate, distinguish between numeric columns and non numeric columns. Numeric columns are related to all the financial variables included in the dataset.

```{r}
data_filtered <- data_filtered %>%
  mutate_at(vars(4:ncol(data_filtered)), as.numeric)
```

Before the last aggregation step, we create a new dataset, called dt, which shows for chartertype the correct qualification of the bank.

```{r}
dt <- as.data.table(data_filtered)

numeric_cols <- names(dt)[4:ncol(dt)]
dt[, (numeric_cols) := lapply(.SD, as.numeric), .SDcols = numeric_cols]
```

```{r}
dt[, RSSD9048 := as.character(RSSD9048)]

# Create a lookup table for chartertype mappings
chartertype_lookup <- c(
  "0" = "Not_known",
  "110" = "Gov_Agency_or_GSE",
  "200" = "Comm_Bank",
  "250" = "Non-dep_Trust",
  "300" = "Sav_Bank",
  "310" = "Sav_loan_Ass",
  "320" = "Coop_Bank",
  "330" = "Cred_Union",
  "340" = "Ind_Bank",
  "400" = "Agreem_Corp",
  "500" = "Holding_Comp",
  "550" = "Insurance",
  "610" = "ESOP/ESOT",
  "700" = "Sec_Broker/Dealer",
  "710" = "Utility_Comp",
  "720" = "Non-Dep_Institution",
  "800" = "Pseudo_entity",
  "810" = "Counting_machine",
  "900" = "Group_of_ind"
)
# Replace numeric values in 'chartertype' with descriptive values using the lookup table
dt[, RSSD9048 := ifelse(is.na(RSSD9048) | RSSD9048 == "0", "Not_known", 
                           chartertype_lookup[RSSD9048])]
```

```{r}
dt_aggregated <- dt[, c(lapply(.SD, sum, na.rm = TRUE), list(RSSD9048 = first(RSSD9048))), 
                    by = .(date, bankid), .SDcols = numeric_cols]

dt_aggregated <- dt_aggregated %>%
  select(date, bankid, RSSD9048, everything())
```

See how many banks are in the dataset

```{r}
unique_bankid_count_agg <- dt_aggregated %>%
  summarise(unique_bankid_count_agg = n_distinct(bankid))

print(unique_bankid_count_agg)
```
The dataset includes a total of 9078 holding banks.

As a further check, we look for duplicate entries, which gives result 0.

```{r}
# Check for multiple lines with the same dateq and bankid
duplicate_entries <- dt_aggregated[, .N, by = .(date, bankid)][N > 1]

# Display the duplicate entries
print(duplicate_entries)
```

```{r}
write.csv(
  dt_aggregated,
  file.path(output_dir, "Final_Dataset_Aggregated.csv"),
  row.names = FALSE
)
```
